\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{multicol}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}
\input{generated_results.tex}
\input{generated_review_results.tex}

\title{Lost in One Index: Measuring How Language and Modality Gaps\\
Compound in English--Indonesian Multimodal Scholarly Search}

\author{Shaban Lubanga\\
Department of Information Systems\\
Institut Teknologi Sepuluh Nopember, Surabaya, Indonesia\\
\texttt{6026251019@student.its.ac.id}}

\date{}

\begin{document}
\maketitle

\begin{abstract}
Scholarly knowledge in Indonesia is distributed across languages. Research
papers are commonly published in English, whereas lectures and theses are
often produced in Indonesian. Retrieval systems that embed these sources in
a single vector index could help connect them, but their cross-lingual
behavior has received limited evaluation. We construct a bilingual,
multimodal test collection containing ten English papers (2{,}720 text
chunks and 67 figures), eight English and eight Indonesian lecture videos
(2{,}101 transcript segments), and 15 translation-paired queries. We then
measure how retrieval changes when the same question is expressed in English
or Indonesian. With an English-only CLIP encoder, paired queries produce
nearly disjoint top-10 lists (mean Jaccard overlap 0.019). Indonesian queries
retrieve English papers in only 3 of 15 cases. The language and modality gaps
also interact, with Indonesian-query-to-figure similarity forming the
lowest-scoring group in the collection (mean 0.498, compared with 0.812 for
same-language transcripts). A multilingual CLIP text encoder improves
consistency substantially (Jaccard 0.474), but favors lecture transcripts
over paper text. Per-group score calibration improves coverage. Figures are
retrieved for up to 7 of 15 queries, and every Indonesian query retrieves an
English paper. However, calibration reduces Indonesian-query effectiveness
(nDCG@10 $0.120 \rightarrow 0.051$), indicating a fairness--precision
tradeoff. Under provisional judgments from a single LLM assessor, BM25
outperforms every dense configuration in both languages, partly because
Indonesian lectures contain code-switched English terminology. A separately
versioned formal Indonesian query set also shows that retrieval is sensitive
to register, although mCLIP is more stable than CLIP or BM25. We release the
collection, provisional judgments, query audit, and code.
\end{abstract}

\section{Introduction}

A student studying deep learning at an Indonesian university often works
across two languages. Much of the primary literature, including arXiv papers
and conference proceedings, is written in English. Explanatory material,
including university lectures, campus theses, and national journals, is
often available in Bahasa Indonesia. A literature review may therefore
require searching both bodies of material. Existing search tools do not
always connect them effectively. An Indonesian question may fail to retrieve
an English paper, while an English query may not surface a relevant
Indonesian lecture.

Multimodal retrieval systems that embed text, figures, and lecture
transcripts into one shared vector space offer a possible solution. If the
representations are well aligned, one query should retrieve across languages
and modalities. Prior work on such systems, including our earlier
prototype~\cite{synthesizer}, has documented a \emph{modality gap}. In a
shared CLIP index, figures score systematically below text and are rarely
retrieved~\cite{liang2022mind}. This paper examines what happens when a
\emph{language} dimension is added, using English--Indonesian as the test
pair. Indonesian is spoken by more than 270 million people, yet
Indonesian-language evaluation resources remain scarce~\cite{nusacrowd,
indonlu}, and to our knowledge no prior work measures cross-lingual behavior
in multimodal scholarly search.

This work makes four contributions.
\begin{enumerate}
  \item \textbf{A bilingual multimodal test collection.} Ten English
  deep-learning papers, eight English and eight Indonesian lecture videos,
  15 translation-paired queries (authored and verified by a native
  speaker), a formally revised Indonesian query variant with a complete
  change audit, and 999 pooled provisional LLM relevance judgments, all
  released publicly.
  \item \textbf{A measurement of the language gap and its interaction with
  the modality gap.} Under English-only CLIP, translation-paired queries
  return nearly disjoint results (Jaccard 0.019). Indonesian queries against
  English figures form the lowest-similarity
  group in the collection.
  \item \textbf{An encoder comparison.} A multilingual CLIP text encoder
  recovers most cross-lingual consistency (Jaccard 0.474) but compresses
  similarity scores and biases retrieval toward spoken-register content,
  displacing paper text.
  \item \textbf{A calibration experiment revealing a fairness--precision
  tradeoff.} Z-normalizing scores within each (modality, language) group
  restores coverage across groups but reduces relevance for Indonesian
  queries by half, showing that balanced exposure and precision are
  competing objectives in a single shared index.
\end{enumerate}

\section{Related Work}

\textbf{Cross-lingual and multilingual retrieval.} Cross-language
information retrieval has traditionally relied on translation~\cite{clir}.
Multilingual sentence encoders such as LaBSE~\cite{labse} and
knowledge-distilled multilingual CLIP~\cite{mclip} enable direct
cross-lingual dense retrieval by aligning languages in one embedding space.
Benchmarks such as MIRACL~\cite{miracl} evaluate multilingual text
retrieval. Indonesian NLP resources are also expanding through efforts such as
IndoNLU~\cite{indonlu} and NusaCrowd~\cite{nusacrowd}. Multimodal
\emph{and} multilingual scholarly retrieval, however, remains largely
unmeasured. Our collection focuses on this intersection.

\textbf{The modality gap.} Liang et al.~\cite{liang2022mind} showed that
contrastively trained multimodal encoders place image and text embeddings
in systematically different regions of the space. In retrieval over
academic content, this can prevent figures from appearing in ranked results.
In our earlier system, figures were never retrieved by unified top-$k$
search over a shared CLIP index~\cite{synthesizer}. The present work measures how this
gap interacts with a language gap and evaluates one mitigation.

\textbf{Multimodal document retrieval.} ColPali~\cite{colpali} retrieves
document pages as images with vision-language embeddings, and
M3DocRAG~\cite{m3docrag} performs retrieval-augmented generation over
multimodal document collections. These systems target visually rich English
documents, whereas our setting adds spoken lectures and a second language.

\textbf{LLM-based relevance judgment.} Recent work shows large language
models can produce relevance judgments that correlate with human assessors
at useful levels~\cite{thomas2024, umbrela}. We adopt this methodology,
with guidelines, per-decision conventions, and known biases documented
alongside the released judgments.

\section{Test Collection and System}

\subsection{Corpus}
The corpus covers a broad range of deep-learning topics for which Indonesian
lecture material is available.

\begin{itemize}
  \item \textbf{Papers (English).} The collection includes ten arXiv papers on
  word2vec~\cite{word2vec}, VGG~\cite{vgg}, ResNet~\cite{resnet},
  Transformer~\cite{transformer}, BERT~\cite{bert}, Mamba~\cite{mamba},
  Jamba~\cite{jamba}, Mamba-360~\cite{mamba360}, Mamba-2~\cite{mamba2},
  and an empirical study of Mamba-based LMs~\cite{empiricalmamba}. These
  papers yield 2{,}720 text chunks and 67 embedded raster figures.
  \item \textbf{Lectures.} The collection includes eight English videos
  (including 3Blue1Brown,
  Stanford CS231N/CS224N, and Michigan lectures) and eight Indonesian
  videos from university and community channels (Machine Learning
  Indonesia, Telkom University course recordings, and others), yielding
  1{,}146 English and 955 Indonesian transcript segments. Indonesian
  transcripts are auto-generated captions, so transcription noise is part of
  the measured condition.
  \item \textbf{Queries.} The evaluation uses 15 study questions in
  translation-paired form
  (English and Indonesian), spanning architectures, training methods, and
  model comparisons. The original Indonesian phrasings were written by the
  author, a native speaker. We preserve them unchanged as the experimental
  baseline and add a separately versioned formal-register variant. The
  revision introduces formal Indonesian equivalents while retaining widely
  used English technical terms in parentheses. Every change and its
  semantic-equivalence assessment is recorded in an audit file.
\end{itemize}

Text and transcripts are chunked to at most 50 words to respect the CLIP
text encoder's 77-token input limit. Figures carry captions parsed from PDF
text or generated by LLaVA-7B~\cite{llava}, and are embedded as the average
of image and caption embeddings. All items carry modality and language
metadata. The index contains 4{,}888 vectors of dimension 512, searched
exactly with FAISS~\cite{faiss} inner product over unit-normalized vectors.

\subsection{Retrieval configurations}
We compare the following configurations at $k{=}10$ over identical corpora.
\begin{itemize}
  \item \textbf{CLIP.} English CLIP ViT-B/32 encodes text and images.
  \item \textbf{mCLIP.} A multilingual text encoder is distilled into the CLIP
  image space~\cite{mclip}, paired with the same image encoder.
  \item \textbf{BM25.} This lexical baseline~\cite{bm25} searches every textual
  representation (chunks, captions, transcripts), both languages mixed.
  \item \textbf{Calibrated variants.} For each dense encoder, similarity
  scores are z-normalized per query \emph{within each (modality, language)
  group} before ranking, so that no group is advantaged by the scale of its
  score distribution.
\end{itemize}

\subsection{Measures}
For every configuration and query language, we report the modality and
content-language distribution of retrieved items. We also report per-query
coverage, including whether a query retrieves at least one figure and at
least one item in each content language. \emph{EN$\leftrightarrow$ID
consistency} is the Jaccard overlap between the result sets for each English
query and its Indonesian translation, averaged across pairs. A
language-consistent system approaches 1. Finally, we report graded
effectiveness (P@10 and nDCG@10) against pooled binary relevance judgments.
The baseline judgments cover 999 items retrieved by the original-query
configurations. A single large language model produced the judgments under
written guidelines. The guidelines apply a strict relevance standard,
distinguish term mentions from relevance, assess meaning regardless of
language, and require visual inspection of figures. We release the
guidelines, decision conventions, and known biases with the data. These
effectiveness values are provisional rather than human-validated evidence,
and the single-assessor design remains a limitation
(Section~\ref{sec:limits}).

For robustness, we repeat every configuration with the formal Indonesian
query variant at $k\in\{5,10,20\}$. We report per-query values, paired
formal-minus-original differences, 95\% bootstrap confidence intervals
(10{,}000 resamples; seed 20260718), and paired standardized effects. Because
new rankings include items outside the original judgment pool, formal-query
effectiveness is reported with judgment coverage and is not treated as
directly comparable to the fully judged baseline.

\section{Results}

\subsection{English-only CLIP separates retrieval by language}
\label{sec:silo}

Table~\ref{tab:dist} shows the modality and language distribution of
retrieved items. Under CLIP, English queries retrieve almost exclusively
English content (148/150 items), while Indonesian queries retrieve almost
exclusively Indonesian content (134/150). Only 3 of 15 Indonesian queries
retrieve any English paper content. Consistency between translation-paired
queries is near zero (Jaccard 0.019, overlap@10 0.033). In other words, the
same question expressed in the other language returns an almost entirely
different result list. Neither query language retrieves figures, which
replicating the modality gap on a corpus $2.3\times$ larger than in our
earlier study~\cite{synthesizer}.

\begin{table}[t]
\centering
\caption{Retrieved items by modality and content language (15 queries
$\times$ top-10). Coverage counts queries returning $\geq$1 figure /
$\geq$1 English-content item / $\geq$1 Indonesian-content item.}
\label{tab:dist}
\small
\begin{tabular}{llrrrrrl}
\toprule
Config & Query lang. & Text & Fig. & Trans. & EN & ID & Coverage (fig/EN/ID)\\
\midrule
CLIP  & EN & 74 & 0 & 76  & 148 & 2   & 0/15 \;\; 15/15 \;\; 2/15\\
CLIP  & ID & 11 & 0 & 139 & 16  & 134 & 0/15 \;\; 3/15 \;\; 15/15\\
mCLIP & EN & 13 & 0 & 137 & 22  & 128 & 0/15 \;\; 8/15 \;\; 15/15\\
mCLIP & ID & 5  & 0 & 145 & 8   & 142 & 0/15 \;\; 5/15 \;\; 15/15\\
BM25  & EN & 85 & 4 & 61  & 146 & 4   & 2/15 \;\; 15/15 \;\; 1/15\\
BM25  & ID & 18 & 2 & 130 & 25  & 125 & 1/15 \;\; 7/15 \;\; 15/15\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Language and modality effects interact}
\label{sec:gaps}

Table~\ref{tab:gap} reports the mean cosine similarity between queries and
\emph{all} indexed items, grouped by (modality, content language). Two
patterns are visible. The first is the \emph{language gap}. Under CLIP,
Indonesian queries score same-language transcripts at 0.812 but English
paper text at only 0.649. This difference of approximately $+0.16$ helps
explain the separation observed in Section~\ref{sec:silo}. The second is the
\emph{modality gap}. Figures score far below all text groups for both query
languages. The effects also interact. The
Indonesian-query$\to$English-figure group (0.498) is the
lowest cell in the table, and its \emph{maximum} (0.622) barely reaches the
\emph{mean} of the competing same-language transcript group. Content that
is both cross-language and cross-modality is doubly disadvantaged.

\begin{table}[t]
\centering
\caption{Query--item cosine similarity by (modality, content language),
mean $\pm$ std over 15 queries $\times$ all items in the group.}
\label{tab:gap}
\small
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{CLIP} & \multicolumn{2}{c}{mCLIP}\\
Group ($n$) & EN queries & ID queries & EN queries & ID queries\\
\midrule
Text EN (2720)       & $.714 \pm .075$ & $.649 \pm .072$ & $.835 \pm .033$ & $.843 \pm .035$\\
Figure EN (67)       & $.556 \pm .043$ & $.498 \pm .041$ & $.615 \pm .034$ & $.617 \pm .033$\\
Transcript EN (1146) & $.761 \pm .049$ & $.703 \pm .050$ & $.860 \pm .024$ & $.877 \pm .026$\\
Transcript ID (955)  & $.718 \pm .051$ & $.812 \pm .056$ & $.880 \pm .023$ & $.899 \pm .027$\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Multilingual encoding reduces the language gap but introduces
register bias}

Under mCLIP the per-group means become nearly identical for English and
Indonesian queries (Table~\ref{tab:gap}, right), and consistency between
paired queries rises from 0.019 to 0.474 Jaccard (overlap@10 0.593). Thus,
changing the encoder removes most of the separation by language. The score
distributions also become narrower, with standard deviations approximately
halved, and their ordering changes. Transcripts in \emph{both} languages
now outscore paper text and account for 137--145 of 150 retrieved items
(Table~\ref{tab:dist}). This shift reduces the representation of papers,
which generally provide the most authoritative content. Figures remain
absent from the results. Multilingual encoding therefore reduces the
language bias but introduces a register bias. More generally, raw
similarity scores are not directly comparable across content groups because
their relative scale depends on the encoder's training distribution.

\subsection{Calibration improves coverage but reduces precision}

Z-normalizing scores within each (modality, language) group forces the
groups into comparable ranges. This substantially changes coverage. Figures
appear for up to 7 of 15 queries, including 5 of 15 Indonesian queries under
calibrated mCLIP. Under CLIP, English-paper coverage for Indonesian queries
increases from 3/15 to 15/15. For Indonesian queries under mCLIP, the number
of transcript items decreases from 145 to 48. Paired-query
consistency also peaks (Jaccard 0.505, overlap@10 0.640, mCLIP-calibrated).

The effectiveness results show a different pattern (Table~\ref{tab:rel}).
Calibration leaves English-query effectiveness approximately unchanged
(CLIP nDCG@10 $.238 \rightarrow .259$; mCLIP $.227 \rightarrow .218$), but
reduces Indonesian-query effectiveness by about half (CLIP
$.120 \rightarrow .051$; mCLIP $.155 \rightarrow .092$). Calibration can
therefore promote items from underrepresented groups even when they are not
relevant to the query. With only 67 figures and a strict relevance standard,
for example, a retrieved figure is often off-topic. These results indicate
that balanced exposure and precision are competing objectives in a single
shared index.

\begin{table}[t]
\centering
\caption{Effectiveness against pooled relevance judgments
(999 items, 127 relevant).}
\label{tab:rel}
\small
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{P@10} & \multicolumn{2}{c}{nDCG@10}\\
Configuration & EN & ID & EN & ID\\
\midrule
BM25              & \textbf{.253} & \textbf{.180} & \textbf{.328} & \textbf{.242}\\
CLIP raw          & .193 & .113 & .238 & .120\\
CLIP calibrated   & .193 & .047 & .259 & .051\\
mCLIP raw         & .160 & .113 & .227 & .155\\
mCLIP calibrated  & .160 & .067 & .218 & .092\\
\bottomrule
\end{tabular}
\end{table}

\subsection{BM25 outperforms all dense configurations under provisional judgments}

BM25 achieves the best effectiveness in both query languages
(Table~\ref{tab:rel}) while retrieving \emph{some} cross-modal content (via
figure captions) and some cross-lingual content. Seven of 15 Indonesian
queries retrieve English material through lexical matching. Code-switching
provides part of this connection because Indonesian technical lectures
often include English terms such as \emph{attention}, \emph{transformer},
and \emph{layer}. These terms provide shared vocabulary across the language
boundary. In this domain-specific corpus, BM25 benefits more from that
shared vocabulary than the dense configurations benefit from semantic
generalization.

\subsection{Retrieval is sensitive to formal Indonesian wording}
\label{sec:robustness}

Table~\ref{tab:robust} compares the frozen original queries with the
separately versioned formal Indonesian set. Formal wording reduces paired
top-10 overlap for every configuration. Raw CLIP falls from
\OriginalClipJaccard{} to \FormalClipJaccard{} Jaccard, and BM25 from 0.071
to \FormalBmTwentyFiveJaccard{}, consistent with both systems benefiting
from English technical tokens in the original code-switched queries. Raw
mCLIP is more stable (\OriginalMclipJaccard{} to \FormalMclipJaccard{};
paired difference \FormalMclipJaccardDifference{}, 95\% bootstrap CI
$[-0.166,0.030]$), while calibrated mCLIP changes by only
\FormalMclipCalibratedJaccardDifference{} (CI $[-0.094,0.062]$). The central
finding is therefore robust to this wording change. Multilingual encoding
provides substantially greater cross-language consistency, although the
absolute size of the gap depends on query register and terminology.

\begin{table}[H]
\centering
\caption{Top-10 EN/ID Jaccard under the original and formal Indonesian query
sets. Judgment coverage is the share of formal Indonesian top-10 items found
in the original LLM-assessed pool; incomplete coverage prevents a controlled
effectiveness comparison.}
\label{tab:robust}
\small
\begin{tabular}{lrrr}
\toprule
Configuration & Original & Formal ID & Judgment coverage\\
\midrule
CLIP raw & \OriginalClipJaccard & \FormalClipJaccard & \FormalClipLabelCoverage\%\\
mCLIP raw & \OriginalMclipJaccard & \FormalMclipJaccard & \FormalMclipLabelCoverage\%\\
BM25 & \OriginalBmTwentyFiveJaccard & \FormalBmTwentyFiveJaccard & \FormalBmTwentyFiveLabelCoverage\%\\
CLIP calibrated & \OriginalClipCalibratedJaccard & \FormalClipCalibratedJaccard & \FormalClipCalibratedLabelCoverage\%\\
mCLIP calibrated & \OriginalMclipCalibratedJaccard & \FormalMclipCalibratedJaccard & \FormalMclipCalibratedLabelCoverage\%\\
\bottomrule
\end{tabular}
\end{table}

\subsection{A bounded end-to-end review evaluation}
\label{sec:review-eval}

The retrieval results determine which evidence is available to the review
generator, but they do not by themselves establish review quality. We
therefore add a bounded end-to-end check using three representative query
pairs: Transformer attention (q01), CNN feature extraction (q06), and Mamba
long-range dependencies (q11). For each English and Indonesian query, we
generate a review from the raw CLIP and raw mCLIP top-10 results using Llama
3.2 at temperature 0. This produces 12 reviews. We measure only properties
that can be checked mechanically: whether source identifiers refer to the
retrieved set, how many retrieved sources are cited, whether all required
sections are present, and the overlap between evidence cited by paired
English and Indonesian reviews. These measures do not assess factual
correctness or human-perceived quality.

All reviews contain the four required sections. Table~\ref{tab:revieweval}
shows that citation identifiers are usually valid, but reviews cite only
16.7--40.0\% of the retrieved sources. More importantly, the evidence cited
by paired English and Indonesian reviews is nearly disjoint. Mean cited-
evidence Jaccard is \ReviewClipEvidenceJaccard{} for CLIP and
\ReviewMclipEvidenceJaccard{} for mCLIP. Multilingual retrieval therefore
does not, in this small sample, guarantee evidence-consistent reviews.

\begin{table}[t]
\centering
\caption{Mechanical review-grounding measures over three query pairs. Values
are percentages except paired evidence Jaccard. A valid identifier names a
retrieved source; it does not prove that the associated claim is supported.}
\label{tab:revieweval}
\small
\begin{tabular}{llrrr}
\toprule
Encoder & Query lang. & Valid IDs & Source coverage & Paired Jaccard\\
\midrule
CLIP  & EN & \ReviewClipEnCitationValidity\% & \ReviewClipEnSourceCoverage\% & \ReviewClipEvidenceJaccard\\
CLIP  & ID & \ReviewClipIdCitationValidity\% & \ReviewClipIdSourceCoverage\% & \ReviewClipEvidenceJaccard\\
mCLIP & EN & \ReviewMclipEnCitationValidity\% & \ReviewMclipEnSourceCoverage\% & \ReviewMclipEvidenceJaccard\\
mCLIP & ID & \ReviewMclipIdCitationValidity\% & \ReviewMclipIdSourceCoverage\% & \ReviewMclipEvidenceJaccard\\
\bottomrule
\end{tabular}
\end{table}

The q11 case illustrates the difference between retrieval and generation.
For the English query, CLIP retrieves three English Mamba lecture segments
in its top three. For the Indonesian query, its top three are Indonesian
transcript segments unrelated to the Mamba architecture, and the resulting
review incorrectly describes Mamba as a classification system. mCLIP places
an English Mamba paper passage second for the Indonesian query and produces
a more relevant overview, although it still makes claims that are not
established by citation syntax alone. The complete reviews, retrieved item
identifiers, and measurements are included in the released artifact.

\section{Discussion}

Our measurements support three practical conclusions. First, uncalibrated
single-index dense retrieval can produce systematically uneven results
across languages and modalities. Retrieval depends not only on relevance,
but also on encoder-induced score differences associated with language,
register, and modality. Second, each mitigation addresses only part of the
problem. Multilingual encoding reduces the language offset but introduces a
register offset, while per-group calibration improves exposure but can
promote irrelevant items from smaller groups. Systems that require both
balanced coverage and precision may benefit from retrieving within each
group and then applying a relevance-aware merge policy. Possible approaches
include learned fusion and lexical--dense hybrids that use code-switched
terms. Third, BM25 remains an important baseline for low-resource-language
deployments and should be included when evaluating multilingual dense
retrievers. The robustness experiment adds an important qualification.
Code-switched terminology can substantially improve lexical retrieval, so
evaluations should report query register rather than treating Indonesian as
a single uniform condition. Finally, the bounded review evaluation shows
that improved multilingual retrieval does not automatically produce
evidence-consistent synthesis. Retrieval and generation should therefore be
evaluated separately.

\section{Limitations}
\label{sec:limits}

The collection is limited to one topic domain, 4{,}888 items, 15 query
pairs, and one language pair. Although the observed differences are large
and consistent, their generalizability to other languages and domains has
not been established. Relevance was judged by a single LLM assessor.
Although the methodology follows recent
practice~\cite{thomas2024,umbrela} and all guidelines, conventions, and
judgments are released for audit, the effectiveness results remain
provisional. Future validation should use at least two bilingual assessors
on a stratified sample spanning query language, modality, configuration,
and relevance label. The validation should report Cohen's $\kappa$ and raw
agreement, then adjudicate disagreements before recalculating P@10 and
nDCG@10. The formal-query runs retrieve previously unjudged items (top-10 judgment
coverage ranges from 52.7\% to 85.3\%), so their apparent effectiveness is
not compared directly with the fully pooled baseline. Indonesian
transcripts are auto-generated captions, so ASR noise and caption quality are confounded
with language effects. We treat this as part of the realistic condition
because a deployed system would index the available captions. Finally,
snippet truncation during judging may undercount the relevance of long text chunks
relative to transcripts.

The review evaluation is intentionally small and structural. It covers only
three query pairs and two raw dense encoders. Valid source identifiers do not
demonstrate factual support, and the qualitative q11 analysis is not a
substitute for systematic review assessment. The experiment establishes a
reproducible diagnostic, not a general estimate of literature-review quality.

\section{Ethics and Artifact Availability}

The corpus is derived from publicly accessible papers and lecture videos.
Copyright remains with the original authors, publishers, and video creators.
The artifact includes derived transcript excerpts and figure crops for audit,
but not pretrained model weights or original source PDFs and videos. Metadata
and reconstruction scripts identify the upstream sources.
Automatically generated Indonesian captions may contain recognition errors,
and LLaVA-generated figure captions may be inaccurate. The relevance
judgments are provisional outputs from a single LLM assessor. Generated
reviews should be checked against cited source material before academic use.

The accompanying artifact contains the versioned queries, query audit,
judgments, result files, generated review samples, and scripts needed to
reproduce the tables. Result files record query and index hashes, model
configuration, platform information, cutoff values, and random seeds. Exact
commands and data-use notes are provided in the artifact documentation. The
project repository is available at
\url{https://github.com/shaban2/Crosslingual_Scholar_Search}. The permanent
Zenodo concept DOI is \url{https://doi.org/10.5281/zenodo.21427042}.

\section{Conclusion}

We presented, to our knowledge, the first measurement of how language and
modality gaps interact in multimodal scholarly retrieval using an
English--Indonesian test collection. An English-only encoder separates
results by language. A multilingual encoder improves alignment but does not
make scores comparable across content groups. Calibration improves coverage
but does not consistently improve relevance. Under provisional judgments
from a single LLM assessor, the lexical baseline achieves the highest
effectiveness. A formal-register query variant confirms the relative
advantage of multilingual encoding while showing that absolute gaps depend
on terminology. The bounded review evaluation further shows that valid
citation identifiers and improved retrieval alignment do not guarantee
evidence-consistent synthesis. We release the collection, provisional
judgments, audited query variants, review samples, and code to support future
work on retrieval and generation that are simultaneously language-fair,
modality-fair, grounded, and precise.

\section*{Acknowledgments}

The retrieval pipeline extends the Multimodal Research Synthesizer, built
jointly with Hippolyte Marc-Antoine Catteau-Verniers, who kindly agreed to
this solo extension. Experiment tooling, analysis scripts, and drafting
were assisted by AI systems (Claude, Anthropic, and Codex, OpenAI). The
author designed the study, verified all results, and takes full
responsibility for the content.

\section*{References}
\begingroup
\renewcommand{\section}[2]{}
\begin{multicols}{2}
\begin{thebibliography}{99}
\scriptsize
\raggedright

\bibitem{synthesizer} S.~Lubanga and H.~M.-A. Catteau-Verniers,
``Multimodal Research Synthesizer,'' software and technical report, 2026.
\url{https://github.com/shaban2/Multimodal-Research-Synthesizer}

\bibitem{liang2022mind} W.~Liang, Y.~Zhang, Y.~Kwon, S.~Yeung, and J.~Zou,
``Mind the Gap: Understanding the Modality Gap in Multi-modal Contrastive
Representation Learning,'' in \emph{Proc.\ NeurIPS}, 2022.

\bibitem{clir} J.-Y. Nie, \emph{Cross-Language Information Retrieval}.
Morgan \& Claypool, 2010.

\bibitem{labse} F.~Feng, Y.~Yang, D.~Cer, N.~Arivazhagan, and W.~Wang,
``Language-agnostic BERT Sentence Embedding,'' in \emph{Proc.\ ACL}, 2022.

\bibitem{mclip} N.~Reimers and I.~Gurevych, ``Making Monolingual Sentence
Embeddings Multilingual using Knowledge Distillation,'' in \emph{Proc.\
EMNLP}, 2020.

\bibitem{miracl} X.~Zhang et al., ``MIRACL: A Multilingual Retrieval
Dataset Covering 18 Diverse Languages,'' \emph{TACL}, vol.~11, 2023.

\bibitem{indonlu} B.~Wilie et al., ``IndoNLU: Benchmark and Resources for
Evaluating Indonesian Natural Language Understanding,'' in \emph{Proc.\
AACL-IJCNLP}, 2020.

\bibitem{nusacrowd} S.~Cahyawijaya et al., ``NusaCrowd: Open Source
Initiative for Indonesian NLP Resources,'' in \emph{Findings of ACL}, 2023.

\bibitem{colpali} M.~Faysse et al., ``ColPali: Efficient Document Retrieval
with Vision Language Models,'' in \emph{Proc.\ ICLR}, 2025.

\bibitem{m3docrag} J.~Cho, D.~Mahata, O.~\.Irsoy, Y.~He, and M.~Bansal,
``M3DocRAG: Multi-modal Retrieval is What You Need for Multi-page
Multi-document Understanding,'' arXiv:2411.04952, 2024.

\bibitem{thomas2024} P.~Thomas, S.~Spielman, N.~Craswell, and B.~Mitra,
``Large Language Models Can Accurately Predict Searcher Preferences,'' in
\emph{Proc.\ SIGIR}, 2024.

\bibitem{umbrela} S.~Upadhyay, R.~Pradeep, N.~Thakur, N.~Craswell, and
J.~Lin, ``UMBRELA: UMbrela is the (Open-Source Reproduction of the) Bing
RELevance Assessor,'' arXiv:2406.06519, 2024.

\bibitem{word2vec} T.~Mikolov, K.~Chen, G.~Corrado, and J.~Dean,
``Efficient Estimation of Word Representations in Vector Space,''
arXiv:1301.3781, 2013.

\bibitem{vgg} K.~Simonyan and A.~Zisserman, ``Very Deep Convolutional
Networks for Large-Scale Image Recognition,'' in \emph{Proc.\ ICLR}, 2015.

\bibitem{resnet} K.~He, X.~Zhang, S.~Ren, and J.~Sun, ``Deep Residual
Learning for Image Recognition,'' in \emph{Proc.\ CVPR}, 2016.

\bibitem{transformer} A.~Vaswani et al., ``Attention Is All You Need,'' in
\emph{Proc.\ NeurIPS}, 2017.

\bibitem{bert} J.~Devlin, M.-W. Chang, K.~Lee, and K.~Toutanova, ``BERT:
Pre-training of Deep Bidirectional Transformers for Language
Understanding,'' in \emph{Proc.\ NAACL}, 2019.

\bibitem{mamba} A.~Gu and T.~Dao, ``Mamba: Linear-Time Sequence Modeling
with Selective State Spaces,'' arXiv:2312.00752, 2023.

\bibitem{jamba} O.~Lieber et al., ``Jamba: A Hybrid Transformer-Mamba
Language Model,'' arXiv:2403.19887, 2024.

\bibitem{mamba360} B.~N. Patro et al., ``Mamba-360: Survey of State Space
Models as Transformer Alternative for Long Sequence Modelling,''
arXiv:2404.16112, 2024.

\bibitem{mamba2} T.~Dao and A.~Gu, ``Transformers are SSMs: Generalized
Models and Efficient Algorithms Through Structured State Space Duality,''
in \emph{Proc.\ ICML}, 2024.

\bibitem{empiricalmamba} R.~Waleffe et al., ``An Empirical Study of
Mamba-based Language Models,'' arXiv:2406.07887, 2024.

\bibitem{llava} H.~Liu, C.~Li, Q.~Wu, and Y.~J. Lee, ``Visual Instruction
Tuning,'' in \emph{Proc.\ NeurIPS}, 2023.

\bibitem{faiss} J.~Johnson, M.~Douze, and H.~J\'egou, ``Billion-scale
Similarity Search with GPUs,'' \emph{IEEE Trans.\ Big Data}, 2019.

\bibitem{bm25} S.~Robertson and H.~Zaragoza, ``The Probabilistic Relevance
Framework: BM25 and Beyond,'' \emph{Foundations and Trends in Information
Retrieval}, vol.~3, no.~4, 2009.

\end{thebibliography}
\end{multicols}
\endgroup

\end{document}
